---
title: "test"
author: "Mehdi Rahmouni"
date: "March 19, 2017"
output: html_document
---

Before starting the exploration, 

```{r, message=FALSE, warning=FALSE}
library('ProjectTemplate')
load.project()
```


```{r, message=FALSE, warning=FALSE}
setwd("C://Users/Dell/Desktop/Dat640/Project/cancer/data")

```

Load libraries

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(caret)
library(randomForest)
library(gbm)
library(rattle)
library(pROC)
library(plyr)
library(e1071)
library(fBasics)
library(Hmisc)
library(base)
library(tidyr)
library(ggplot2)
```

use the variable wbcd to manipulate the dataset

```{r, message=FALSE, warning=FALSE}
wbcd <- read.csv("C://Users/Dell/Desktop/Dat640/Project/cancer/data/breast_cancer_wisconsin.csv",
                 header = TRUE)
```                 


#Data exploration

check the dataset structure, variables types

```{r, echo=TRUE}
str(wbcd)
```

column names

```{r, echo=TRUE}
names(wbcd)
```

number of rows

```{r}
nrow(wbcd)
```

number of columns

```{r}
ncol(wbcd)
```

top six rows

```{r message = FALSE, warning = FALSE}
head((wbcd),6)
```

last six rows

```{r message = FALSE, warning = FALSE}
tail((wbcd),6)
```

outcome distribution

```{r}
table(wbcd$diagnosis)
```


#Exploratory data analysis

number of rows and columns
 
```{r}
dim(wbcd)
```

6 number summary
 
```{r}
summary(wbcd)
```

find which column contains missing values

```{r}
colnames(wbcd)[colSums(is.na(wbcd)) > 0] 
```

skewness
```{r warning= FALSE, message= FALSE}
sapply(wbcd,skewness) 
```

kurtosis
```{r warning=FALSE, message=FALSE}
sapply(wbcd,kurtosis)
```

histogram plots

```{r}
wbcd <- wbcd[, -1] %>%
  gather(measure, value, radius_mean:fractal_dimension_worst)

ggplot(data = wbcd, aes(x = value, fill = diagnosis, color = diagnosis)) +
  geom_density(alpha = 0.3, size = 1) +
  geom_rug() +
  scale_fill_brewer(palette = "Set2") +
  scale_color_brewer(palette = "Set2") +
  facet_wrap( ~ measure, scales = "free_y", ncol = 3)
```


Print SessionInfo(), to enable reproducibility

```{r echo = FALSE, message = FALSE, warning = FALSE}
sessionInfo()
```