---
title: "Data mining using Wisconsin breast cancer dataset"
author: "Mehdi Rahmouni"
date: "January 23, 2017"
output:
  html_document: default
 ---

#Load ProjectTemplate() library


First we need to load the project which will automatically load the dataset. 


```{r, message=FALSE, warning=FALSE}
library('ProjectTemplate')
load.project()
```


```{r, message=FALSE, warning=FALSE}
setwd("C://Users/Dell/Desktop/Dat640/Project/cancer/data")

```

Load libraries

```{r echo=FALSE, message=FALSE, warning=FALSE}
if (!require(caret)) install.packages("caret",dependencies=TRUE)
if (!require(randomForest)) install.packages("randomForest",dependencies=TRUE)
if (!require(gbm)) install.packages("gbm",dependencies=TRUE)
if (!require(rattle)) install.packages("rattle",dependencies=TRUE)
if (!require(pROC)) install.packages("pROC",dependencies=TRUE)
if (!require(plyr)) install.packages("plyr",dependencies=TRUE)
if (!require(e1071)) install.packages("e1071",dependencies=TRUE)
if (!require(fBasics)) install.packages("fBasics",dependencies=TRUE)
if (!require(Hmisc)) install.packages("Hmisc",dependencies=TRUE)
if (!require(base)) install.packages("base",dependencies=TRUE)
if (!require(tidyr)) install.packages("tidyr",dependencies=TRUE)
if (!require(ggplot2)) install.packages("ggplot2",dependencies=TRUE)
if (!require(reshape2)) install.packages("reshape2",dependencies=TRUE)

require(caret)
require(randomForest)
require(gbm)
require(rattle)
require(pROC)
require(plyr)
require(e1071)
require(fBasics)
require(Hmisc)
require(base)
require(tidyr)
require(ggplot2)
require(reshape2)
```

use the variable wbcd to manipulate the dataset

```{r, message=FALSE, warning=FALSE}
wbcd <- read.csv("C://Users/Dell/Desktop/Dat640/Project/cancer/data/breast_cancer_wisconsin.csv",
                 header = TRUE)
```                 


#Data exploration

check the dataset structure, variables types

```{r, echo=TRUE}
str(wbcd)
```

find column names

```{r, echo=TRUE}
names(wbcd)
```

dataset number of rows

```{r}
nrow(wbcd)
```

dataset number of columns

```{r}
ncol(wbcd)
```

top six rows

```{r message = FALSE, warning = FALSE}
head((wbcd),6)
```

last six rows

```{r message = FALSE, warning = FALSE}
tail((wbcd),6)
```

outcome distribution

```{r}
table(wbcd$diagnosis)
```


#Exploratory data analysis

number of rows and columns
 
```{r}
dim(wbcd)
```

6 number summary
 
```{r}
summary(wbcd)
```

find which column contains missing values

```{r}
colnames(wbcd)[colSums(is.na(wbcd)) > 0] 
```

skewness

```{r warning= FALSE, message= FALSE}
sapply(wbcd,skewness) 
```

kurtosis

```{r warning=FALSE, message=FALSE}
sapply(wbcd,kurtosis)
```

data visualization using histogram plots

```{r echo=FALSE, warning=FALSE, message=FALSE, out.width = '\\maxwidth'}
wbcd_hist <- melt(wbcd[,-c(1,33)])

ggplot(wbcd_hist,aes(x = value, fill = diagnosis, color = diagnosis)) + 
  facet_wrap(~variable,scales = "free_x") + 
  geom_histogram(bins = 10)
```

#Data transformation

load the dataset again, because after the plot we have a completly different dataset 


```{r}
wbcd <- read.csv("C://Users/Dell/Desktop/Dat640/Project/cancer/data/breast_cancer_wisconsin.csv",
                 header = TRUE)
```

impute id, X 

```{r}
wbcd$id <- NULL
wbcd$X <- NULL
```

assign Benign, Malignant to diagnosis and change type to factor

```{r}
wbcd$diagnosis <- factor(wbcd$diagnosis, levels = c("B", "M"),
                         labels = c("Benign", "Malignant"))
```

verify that id and X are gone

```{r}
str(wbcd)
```

check number of observations and percentage

```{r}
cbind(Frequency = table(wbcd$diagnosis), Percentage = (prop.table(table(wbcd$diagnosis))* 100))
```

data summary

```{r}
summary(wbcd)
```

#Model building

ensure reproducibility with set.seed()

```{r}
set.seed(1)
```

split the dataset into training/tersting (70% training, 30% testing)

```{r}
inTrain <- createDataPartition(wbcd$diagnosis, p=0.7, list=FALSE)
training <- wbcd[inTrain,]
testing <- wbcd[-inTrain,]
```

training/testing dimensions

```{r}
dim(training)
dim(testing)
```

outcome distribution in the testing set

```{r}
table(testing$diagnosis)
```

build a random forest model using the caret package

```{r}
set.seed(1)
rf_model <- train(diagnosis ~ ., data=training, method="rf")
rf_model
```

```{r}
rf_model$finalModel
```

performance on testing test

```{r}
pred_test <- predict(rf_model, testing)
confusionMatrix(pred_test, testing$diagnosis, positive = "Malignant")
```

#Model tuning 

tune Rf model to find best mtry
using cross-validation

```{r}
ctrl <- trainControl(method="repeatedcv", repeats=3)
grid <- expand.grid(mtry = c(1, 2, 3, 5, 7, 10, 15, 20, 30))
set.seed(1)
rf_tune_model <- train(diagnosis ~ ., data=training, method="rf", tuneGrid = grid, trControl = ctrl)
rf_tune_model
```

```{r}
rf_tune_model$finalModel
```

plot the new (tuned) mtry value

```{r}
plot(rf_tune_model)
```

performance of updated model on testing set

```{r}
pred_tuned_test <- predict(rf_tune_model, testing)
confusionMatrix(pred_tuned_test, testing$diagnosis, positive="Malignant")
```

#Other performance measures (Variable importance, ROC curve)

variable importance plot

```{r}
rf_var <- varImp(rf_tune_model)
plot(rf_var)
```

ROC curve

```{r}
rf_roc <- roc(testing$diagnosis,
              predict(rf_tune_model, testing, type = "prob")[,"Malignant"],
              levels = rev(levels(testing$diagnosis)))
roc_value <- auc(rf_roc)

plot.roc(smooth(rf_roc), main = "RF ROC Curve", col = "blue")
legend("bottom", legend = roc_value, col=c("#1c61b6"), lwd=2)

```


print SessionInfo(), to enable reproducibility

```{r echo = FALSE, message = FALSE, warning = FALSE}
sessionInfo()
```