---
title: "Data mining using Wisconsin breast cancer dataset"
author: "Mehdi Rahmouni"
date: "January 23, 2017"
output: html_document
---

##Load libraries
```{r}
library(knitr)
```


##Print SessionInfo(), to enable reproducibility
```{r echo = FALSE, message = FALSE, warning = FALSE}
sessionInfo()
```


## Seting up working directory


Before we start our data exploration,                                          
we need to set up our working directory


```{r message = FALSE, warning = FALSE}
setwd("C://Users//Dell/Desktop/DAT640/Project/")
```

## Data exploration

In this step we start by importing our dataset,                                
and saving it into a dataframe using read.csv command

```{r message = FALSE, warning = FALSE}
wisconsin_bcd <- read.csv("breast_cancer_wisconsin.csv", stringsAsFactors = FALSE)
```

Still in data exploration                                                   
we start by checking the structure of the dataset                           
using the str command 

```{r message = FALSE, warning = FALSE}
str(wisconsin_bcd)
```


## Data exploration

```{r message = FALSE, warning = FALSE}
head((wisconsin_bcd),6)
```
